---
title: "AR_assignment_2_Staub"
author: "Alexander Staub"
date: "July 9, 2020"
output: 
  html_document:
    includes:
      after_body: footer.html
---

## Question 1 
How are the observed firm characteristics distributed across the presence and absence of the treatment? What does this suggest?

### Answer 1

first, create a table of summary statistics with the psych package of the firm characteristic variables x1 to x4. Check whether at least one variable's mean is significantly different in treatment group compared to control

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(psych)
library(tidyverse)
library(readxl)

#load data
df_raw <- read_excel("C:/R work/applied econometrics/applied_econometrics/Data/Session1data.xlsx", 
                     sheet = 1)

#compare summary statistics
psych::describeBy(df_raw[,c(2:5)], group = df_raw$D)

#do t_test for mean differences
t.test(df_raw$x4~df_raw$D)
```

calculating the standardized difference to determine which variables show a significant imbalance between control  and treatment as well as whether at least one of the variables is creating imbalance with an omnibus Xi-squared test (see Hansen and Bowers, 2008)

```{r warning=FALSE, message=FALSE}
library(RItools)
#indices of covariance imbalances
treated<-(df_raw$D==1)
cov<-df_raw[,2:5]
std.diff<-apply(cov,2,function(x) 100*(mean(x[treated])- mean(x[!treated]))/(sqrt(0.5*(var(x[treated])+var(x[!treated])))))
abs(std.diff)
#check whether at least one of the variables is causing statistical imbalance with a chi-squared test
xBalance(D ~ x1 + x2 + x3 + x4, data = df_raw, report = c("chisquare.test"))
```


As one can see in the ouput above, means differ among the firm characteristic variables, however not dramatically. looking at the standardized differences, no absolute scores are higher than 25% (set as suspect limit by Stuart and Rubin, 2008) and the chi squared test does not imply significance at the 0.95 significance level

## Question 2

Use propensity scores to match the presence of the treatment (‘treatment group’) to the absence of the treatment (‘control group’) based on the observed firm characteristics (Note: make sure that the maximum difference in the probability of treatment between the treatment and matched control is less than or equal to 1 percentage point)

i. How are the observed firm characteristics distributed across the presence and absence of the treatment for this matched sample?

ii. What is the difference in the outcome variable Y between the presence and absence of the treatment for the propensity-matched sample?

### Answer 2

First calculate the propensity score for the two groups with a logistic regression using the firm characteristic variables as matching variables. As I am not able to find any R documentation on setting a minimum distance in propensity score for matching individuals, I will run 3 matching procedures and use the matching procedure which provides the best support. All are estimated without replacement and 1:1
 1. Nearest Neighbour greedy
 2. Optimal
 3. Nearest Neighbour Caliper, setting caliper at 0.25 as this seems to be the standard (Rosenbaum and Robin, 1985)

```{r warning=FALSE, message=FALSE}
library(MatchIt)
library(stats)

#matching with nearest neighbour, greedy, no replacement
matched_nn_df <- matchit(D~ x1 + x2 + x3 + x4, method = "nearest", data = df_raw)

#matching with nearest neighbour, no replacement
matched_opt_df <- matchit(D ~ x1 + x2 + x3 + x4, method = "optimal", data = df_raw)

#matching with nearest neighbour caliper, no replacement
matched_calip_df <- matchit(D ~ x1 + x2 + x3 + x4, method ="nearest", caliper = 0.25, data = df_raw)


```

based on the mean difference and qq Max values after the different matching procedures, optimal matching was chosen as the best method for reducing mean difference (check appendix item A to inspect the summary tables for matching procedure).

below find plots of the changes in propensity score distribution before and after using optimal matching procedure

```{r}
plot(matched_opt_df, type = "hist")
```

i) below, find the covariate distribution in presence and absence of treatment with the optimally matched data

```{r}
#get the matched data
df_opt_match <- match.data(matched_opt_df)

#run the summary of the matched data
psych::describeBy(df_opt_match[,2:5], group = df_opt_match$D)
```

after matching, the distribution has improved markedly across all covariates comparing absence and presence of treatment, apart for covariate x3. 

ii) find below a t-test to compare the difference in the outcome variable with the matched data

```{r}
#t test with matched data
t.test(df_opt_match$y ~ df_opt_match$D, conf.level = 0.95)
#unmatched
t.test(df_raw$y ~ df_raw$D, conf.level = 0.95)
```

according to the t-test above, the difference in mean outcomes with the matched data is not significant at the 0.95 confidence level, while there is a signficant difference at the 0.95 level for the unmatched data. 

## Quesiton 3
Use propensity scores to create a sample that is on “common support” and estimate the treatment effect using a regression with covariates.

### answer 3

run a matching procedure that establishes common support. Herein, nearest neighbour greedy, 1:1 without replacement, discarding observations that fall outside the area of common support is implemented

```{r}
#estimate the model around common support
matched_nn_cs_df <- matchit(D ~ x1 + x2 + x3 + x4, data = df_raw, discard="both", type="nearest", m.order="largest")

#save the matched data
df_nn_cs_match <- match.data(matched_nn_cs_df)
```

restricting the matching procedures to create a sample that is on common support leads to the discarding of two observations in the control group

The treatment effect is estimated below

```{r}
summary(lm(y~D+x1+x2+x3+x4, data = df_nn_cs_match))
```
estimating the treatment effect with the matched sample leads to an insignificant treatment effect

## Question 4

Use propensity scores to weight observations and estimate the average treatment effect (ATE) and the treatment on the treated (TOT).

### answer 4

find below the average treatment effect on the treated using propensity score weighting
```{r warning=FALSE, message=FALSE}
if (!require(survey)) install.packages("survey"); library(survey)  

#trying weightit package
if (!require(WeightIt)) install.packages("WeightIt"); library(WeightIt)

W.out <- weightit(D ~ x1 + x2 + x3 + x4,
        data = df_raw, estimand = "ATT", method = "ps")

#check the summary
summary(W.out)
#get the weights
weights <- svydesign(ids = ~1, weights = W.out$weights, data = df_raw)

#run weighted regression
model_w_att <- svyglm(y ~ D + x1 + x2 + x3 + x4, design = weights)
summary(model_w_att)

```

using the weighting procedure, the treatement effect on the treated is significant at the 0.95 significance level

below the weighting procedure for the average treatment effect on the treated
```{r}
#estimate the propensity scores
ps_ate <- glm(D ~ x1 + x2 + x3 + x4, data = df_raw, family = "binomial")
#attach ps scores to the dataset
df_raw$psvalues <- predict(ps_ate, type = "response")
#extract the weights
df_raw$w_ate <- ifelse(df_raw$D == 1, 1/df_raw$psvalues, 1/(1-df_raw$psvalues))
#calculate the model
model_w_ate <- lm(y ~ D + x1 + x2 + x3 + x4, data = df_raw, weights = (w_ate))
summary(model_w_ate)
```
The ATE using propensity score weights is also significant at the 0.9 5significance level

In summary, using matching leads to a lower adjusted R-sqaured as well as an insignificant treatment effect (at least at 0.95 level) as well as a lower adjusted R-squared

## Question 5
Use entropy balancing to estimate the treatment effect of D on the outcome Y?

*  i. Use as balance constraints the first moment (mean).

*  5.i.1. How are the observed firm characteristics distributed across the presence and absence of the treatment for this matched sample?

*  5.i.2. What is the difference in the outcome variable Y between the presence and absence of the treatment for the propensity-matched sample?

*  ii. Use as balance constraints the first and second moment (mean and variance).

*  5.ii.1. How are the observed firm characteristics distributed across the presence and absence of the treatment for this matched sample

*  5.ii.2. What is the difference in the outcome variable Y between the presence and absence of the treatment for the propensity-matched sample?


### answer 5

5.i.1 run the entropy balancing and check the balance of the observed firm characteristics
```{r warning=FALSE, message=FALSE}
if (!require(ebal)) install.packages("ebal"); library(ebal)
library(cobalt)
#run the entropy balancing on the first moment
W.out.ebal.1 <- weightit(D ~ x1 + x2 + x3 + x4,
        data = df_raw, estimand = "ATT", method = "ebal", moments = 1)

#check the balance
bal.tab(W.out.ebal.1, m.threshold = 0.05, disp.v.ratio = T)

```
one can see that the entropy balancing has yielded very good balance across the board

5.i.2 estimate the model with the reweighted data
```{r}
#get the weights
weights.ebal.1 <- svydesign(ids = ~1, weights = W.out.ebal.1$weights, data = df_raw)

#run weighted regression
model_w_ebal_1 <- svyglm(y ~ D + x1 + x2 + x3 + x4, design = weights.ebal.1)
summary(model_w_ebal_1)
```
using the entropy balancing method fixed on 1st moment, the difference in outcome between treatment and non-treatment ist 0.07 at the 0.95 significance level


5.i.2 estimate the weights for entropy balancing across 1st and 2nd moment

```{r}
#run the entropy balancing on the first moment
W.out.ebal.2 <- weightit(D ~ x1 + x2 + x3 + x4,
        data = df_raw, estimand = "ATT", method = "ebal", moments = 2)

#check the balance
bal.tab(W.out.ebal.2, m.threshold = 0.05, disp.v.ratio = T)
```
as in the prior example, the entropy balancing yields good balancing

5.ii.2 estimate the model with the reweighted data
```{r}
#get the weights
weights.ebal.2 <- svydesign(ids = ~1, weights = W.out.ebal.2$weights, data = df_raw)

#run weighted regression
model_w_ebal_2 <- svyglm(y ~ D + x1 + x2 + x3 + x4, design = weights.ebal.2)
summary(model_w_ebal_2)
```
the results for using the 1st and 2nd moment yield the same results, however it may be the case that the entropy balancing moment selection in the package I am using is not doing what I expect it to do. However, I am not able to find any reliable documentation on an R package for entropy balancing to solve the problem set as set out in this assignment. 

## Question 6
Provide an overview table with all of the above treatment effects, as well as the treatment effect of the “regular regression” that was part of the Session 1 assignment.

## answer 6

*  Treatment effect A2 assignement 1: 0.0697 at sig level 0.9
*  Treatment effect A3 assignement 1: (w. industry): 0.0943 at sig level 0.9
*  Treatment effect A4 assignement 1: (w. clustered SE): 0.0943 insignificant

*  Treatment effect A3 assignemnet 2: (w. matching and common support): 0.0690 at sig level 0.9
*  ATT effect A4 assignment 2: (w. ps reweighting): 0.0686 at sig level 0.95
*  ATE effect A4 assignment 2: (w. ps reweighting): 0.0746 at sig level 0.95
*  Treatment effect A5 assignment 2: (entropy balancing 1st moment): 0.0684 at sig level 0.95
*  Treatment effect A5 assignemnet 2: (entropy balancing 2nd moment): 0.0684 at sig level 0.95





## Appendix

Appendix A: Summaries of the matching procdeures in question 2

nearest neighbour greedy
```{r}
summary(matched_nn_df, standardize = T)
```

optimal matching
```{r}
summary(matched_opt_df, standardize = T)
```

nearest neighbour caliper
```{r}
summary(matched_calip_df, standardize = T)
```

Appendix B: Summary of nearest neighbour matching, discarding elements outside of common support

```{r}
summary(matched_nn_cs_df)
```

